\begin{abstract}
  We present a criterion for checking local and global deadlock freedom of finite state systems expressed in BIP: a component-based framework for the
  construction of complex distributed systems.  Our criterion is evaluated by model-checking a set of subsystems of the overall large system. If
  satisfied in small subsystems, it implies deadlock-freedom of the overall system. If not satisfied, then we re-evaluate over larger subsystems,
  which improves the accuracy of the check.  When the subsystem being checked becomes the entire system, our criterion becomes complete for
  deadlock-freedom.  Hence our criterion can only fail to decide deadlock-freedom because of computational limitations: state-space explosion sets in
  when the subsystems being checked become too large.  Our method thus combines the possibility of fast response together with theoretical
  completeness.  Other criteria for deadlock-freedom, in contrast, are incomplete in principle, and so may fail to decide deadlock-freedom even if
  unlimited computational resources are available.  Also, our criterion certifies freedom from local deadlock, in which a subsystem is deadlocked
  while the rest of the system executes. Other criteria only certify freedom from global deadlock. We present experimental results for dining
  philosophers and for a multi token-based resource allocation system, which subsumes several data arbiters and schedulers, including Milner's token
  based scheduler.
\end{abstract}



