


%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The supercycle formation condition}

We use the structural properties of supercycles (\secn{supercycle-structural}) and the dynamics of wait-for graphs (\prop{wait-for-edge-preservation})
to define a condition that must hold whenever a supercycle is created. Negating this condition then implies the absence of supercycles.


\begin{proposition}[Supercycle formation condition] \label{prop:supercycle-formation}
Assume that $s \goesto[\act] t$ is a transition of $(\B, Q_0)$, $\wfg{\B}{s}$ is supercycle-free, and that $\wfg{\B}{t}$
contains a supercycle.  Then, in $\wfg{\B}{t}$, there exists a $\CC$ such that
\bn
\item $\CC$ is a subgraph of $\wfg{\B}{t}$, \ie $\CC \subg \wfg{\B}{t}$
\item $\CC$ is strongly connected
\item $\CC$ is a supercycle
\item  in $\wfg{\B}{t}$, there is no wait-for edge from a node in $\CC$ to a node outside of $\CC$
\item there exists a component $\B_i \in \cmps{\act}$ such that $\B_i$ is in $\CC$
\en
\end{proposition}
%
\begin{proof}
  By assumption, there is a supercycle $\SC$ that is a subgraph of $\wfg{\B}{t}$.  By Proposition~\ref{prop:supercycle:contains-mssc}, $\SC$ contains
  a subgraph $\CC$ that is strongly connected, is itself a supercycle, and such that there is no wait-for-edge from a node in $\CC$ to a node outside
  of $\CC$.  This establishes Clauses~1--4.

Now suppose $\B_i \not\in \CC$ for every $\B_i \in \cmps{\act}$. Then, no edge in $\CC$ is
$\B_i$-incident.  Hence, by Proposition~\ref{prop:wait-for-edge-preservation}, every edge in $\CC$
is an edge in $\wfg{\B}{s}$. Hence $\CC$ is a subgraph of $\wfg{\B}{s}$.
%
Now let $v$ be an arbitrary node in $\CC$.
%
Suppose $v$ is a component $\B_j$.  By assumption, $\B_j \not\in \cmps{\act}$, and so
$s \pj \B_j = t \pj \B_j$ by Definition~\ref{def.bip.composition}. Hence $\B_j$ enables the same set
of interactions in state $s$ as in state $t$. Also, in $\wfg{\B}{t}$, all of $\B_j$'s wait-for edges
must end in an interaction that is in $\CC$, since $\CC$ is a supercycle in $\wfg{\B}{t}$. Hence the
same holds in $\wfg{\B}{s}$.
%
If $v$ is an interaction, it must also have a wait-for-edge $e'$ to some component $\B_j \in \CC$,
since $\CC$ is a supercycle in $\wfg{\B}{t}$. Hence this also holds in $\wfg{\B}{s}$.
%
Hence $v$ has enough successors in $\CC$ to satisfy the supercycle definition (\defn{supercycle}).
%
We conclude that $\CC$ by itself is a supercycle in $\wfg{\B}{s}$, which contradicts the assumption
that $\wfg{\B}{s}$ is supercycle-free. Hence, $\B_i \in \CC$ for some $\B_i \in \cmps{\act}$, and so
Clause~5 is established.  
\end {proof}


%% overview of the four criteria: intrroduce strong connectedness violation

The supercycle formation condition (Proposition~\ref{prop:supercycle-formation}) tells us that, when a supercycle $\SC$ is created, some component
$\B_i$ that participates in the interaction $\act$ whose execution created $\SC$, must be a node of a strongly connected component $\CC$ of $\SC$, and
moreover $\CC$ is itself a supercycle in its own right. In a sense, $\CC$ is the ``essential'' part of $\SC$.  We use this to formulate a condition
that prevents the formation of supercycles. For transition $s \goesto[\act] t$, we determine for every component $\B_i \in \cmps{\act}$ whether it is
possible for $\B_i$ to be a node in a strongly-connected supercycle $\CC$ in $\wfg{\B}{t}$.  There are two ways for $\B_i$ to not be a node in a
strongly-connected supercycle: 
%
\bn
\item \label{conditions:no-sc-member}   \textit{no supercycle membership}: $\B_i$ is not a node of any supercycle, \ie $\neg \scyc{\B}{s}{\B_i}$.
\item  \label{conditions:no-strongly-connected} \textit{no strong-connectedness}: $\B_i$ is a node in a supercycle, but not a node in a \emph{strongly-connected} supercycle. 
\en
%
Hence, for a BIP system $(\B, Q_0)$, our fundamental criterion for the prevention of supercycles is that, for every reachable transition
$s \goesto[\act] t$ resulting from execution of $\act$, in the resulting state $t$, every component $\B_i$ of $\act$ must violate at least one of the
above two conditions.  Condition~(\ref{conditions:no-sc-member}) is just supercycle violation, as in \defn{supercycle.violation}.
Condition~(\ref{conditions:no-strongly-connected}) is violation with respect to a strongly connnected supercycle, \ie non-membership in a strongly
connected supercycle. Technically, this is implied by supercycle violation, and so the disjunction of the two conditions is equivalent to 
Condition~(\ref{conditions:no-sc-member}). It is however, convenient to use the disjunction of the two conditions, since
we will formulate local versions of these violation conditions, and the implication does not necessarily hold for the local versions.

For a given BIP system $(\B, Q_0)$ and interaction $\act$, we use $\GAO(\B, Q_0, \act)$ to denote the deadlock-freedom criterion based on the
disjunction of Conditions~(\ref{conditions:no-sc-member}) and (\ref{conditions:no-strongly-connected}) above. 
This criterion is, in a sense, the ``most general'' criterion for supercycle-freedom.
%
If $\GAO(\B, Q_0, \act)$ holds, and global state $s$ is supercycle-free, and $s \goesto[\act] t$, then it follows (as we establish below) that global
state $t$ is also supercycle-free.  So, by requiring (1) that all initial states are supercycle-free, and (2) that $\GAO(\B, Q_0, \act)$ holds for all
interactions $\act \in \gamma$, we obtain, by straightforward induction on length of executions, that every reachable state is supercycle-free.

It also follows that any condition which implies $\GAO(\B, Q_0, \act)$ is also sufficient to guarantee  supercycle-freedom, and
hence deadlock-freedom. We exploit this in two ways:
\bn

\item To define a ``linear'' condition, $\GLin$, that is easier to evaluate than $\GAO$, since it requires only the
evaluation of lengths of wait-for-paths, \ie it does not have the ``alternating'' character of $\GAO$.  It also implies $\GAO$.

\item To define ``local variants'' of $\GAO$ and $\GLin$, which we denote as $\LAO$ and $\LLin$, respectively.  $\LAO$ and $\LLin$ can be evaluated in
  small subsystems of $(\B, Q_0)$. When either $\LAO$ or $\LLin$ holds in a small subsystem, we confirm deadlock freedom of $(\B, Q_0)$ without
  state-explosion. The local conditions imply the corresponding global ones, \ie they are sufficient but not necessary for deadlock-freedom.

\en
%
We therefore now have four deadlock-freedom conditions: $\GAO$ (global alternating), $\LAO$ (local alternating), $\GLin$ (global linear), and $\LLin$
(local linear). There are all concrete instances of the abstract version of the deadlock-freedom condition given in \secn{abstract-scfree-conditions}.





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{A Global AND-OR Condition for Deadlock Freedom}
   \label{s:global.ANDOR}
%   \input{globANDORCond}

%\subsection{General supercycle violation condition}

In this section, the operators $\SFs, \VFs$ depend on the global state
$t$, since we wish to show that the transition $s \goesto[\act] t$
does not create a supercycle in state $t$.
%
Towards this end, we first formalize violation of strong connectedness
(Condition~\ref{conditions:no-strongly-connected} above) as follows.

\begin{definition}[Strong connectedness violation, $\connViol{v}{t}$]
\label{def:sConn.violation}
 Let $v$ be a node of $\wfg{\B}{t}$.   Then $\connViol{v}{t}$ holds iff there does not exist a 
strongly connected supercycle $\SSC$ such that $v \in \SSC$ and $\SSC \subg \wfg{\B}{t}$.
\end{definition}



The general supercycle violation condition is then a disjunction of the supercycle violation condition
and the strong connectedness violation conditions.


\begin{definition}[General supercycle violation, $\genViol{v}{t}$]
\label{def:formation.violation} 
\label{defn:formation.violation} 
Let $v$ be a node of $\wfg{\B}{t}$.
Then 
$\genViol{v}{t}  \df \scV{v}{t}  \lor \connViol{v}{t}$.
%$\genViol{v}{t}  \df (\exs d \ge 1: \scViol{v}{d}{t}) \lor \connViol{v}{t}$.
\end{definition}
%
Let $s \goesto[\act] t$ be a reachable transition. If, for every $\B_i \in \cmps{\act}$, $\formViol{v}{t}$ holds, then, as we show below,
$s \goesto[\act] t$ does not introduce a supercycle, \ie if $s$ is supercycle-free, then so is $t$. As stated above, we formulate below a ``local''
version of the general condition, which can be evaluated in ``small subsystems'', and so we often avoid state-explosion.
%
% THIS IS GIVEN IN PREVIOUS SECTION
% However, evaluating this condition over all global transitions is subject to state explosion, and so we formulate below a ``local'' version of the
% general condition, which can be evaluated in ``small subsystems'', and so we often avoid state-explosion. Hence the advantage of the local versions is
% that they are usually efficiently computable, as we show in the sequel.  We also formulate a ``linear'' condition (both global and local), which is
% simpler (but ``more incomplete'') than the general condition, and so is easier to evaluate.
%
% THIS IS GIVEN IN PREVIOUS SECTION, WITHOUT THE FORMALISM
We reitrate that $\scV{v}{t}$ implies that $v$ cannot be in a supercycle. Hence, $v$ cannot be in a strongly-connected supercycle.
Hence $\scV{v}{t} \imp \connViol{v}{t}$, so that 
$\scV{v}{t}  \lor \connViol{v}{t} \ev  \connViol{v}{t}$.
We give the formation violation condition in this manner, since 
the implication does not necessarily hold for the local versions of $\scV{v}{t}$ and $\connViol{v}{t}$.

This discussion leads to the formal definition of $\GAO$: 
after execution of interaction $\act$, all $\B_i \in \cmps{\act}$ exhibit a general supercycle-violation, as given by $\genViol{\B_i}{t}$ above.

\begin{definition}[$\GAO(\B, Q_0, \act)$] \label{def:global.ANDOR-cond} \label{defn:global.ANDOR-cond} 
Let $s \goesto[\act] t$ be a reachable transition of $(\B, Q_0)$.
Then, for every component $\B_i \in \comps{\act}$, the formation violation condition holds in state $t$.
Formally,\\
\ind  $\fas \B_i \in \comps{\act}, \genViol{\B_i}{t}$.
\end{definition}


\begin{theorem} \label{thm:GAO.SC-free-preserving}
$\GAO$ is supercycle-freedom preserving.
\end{theorem}
%
\begin{proof}
We establish:
for every reachable transition $s \la{\act} t$,
$\wfg{\B}{s}$ is supercycle-free implies that $\wfg{\B}{t}$ is
supercycle-free. Our proof is by contradiction, so we assume the existence of a reachable transition
$s \la{\act} t$ such that $\wfg{\B}{s}$ is supercycle-free and $\wfg{\B}{t}$ contains a supercycle.
%
By Proposition~\ref{prop:supercycle-formation}
 there exists a component $\B_i \in \cmps{\act}$ such that $\B_i$ is in $\CC$, where 
$\CC$ is a strongly connected supercycle that is a subgraph of $\wfg{\B}{t}$.
%
Since $\CC$ is a strongly connected supercycle, we have,
 by Definition~\ref{def:sConn.violation}, that $\neg \connViol{\B_i}{t}$ holds.
%
Since $\CC$ is a supercycle, we have, by Proposition~\ref{prop:scViol-iff-notInSC}, 
that $\neg \scV{\B_i}{t}$ holds. %$\neg (\exs d \ge 1: \scViol{\B_i}{d}{t})$ holds.
%
Hence, by Definition~\ref{def:formation.violation}, $\neg \formViol{\B_i}{t}$.
But, by Definition~\ref{def:global.ANDOR-cond}, we have $\formViol{\B_i}{t}$.
Hence, we have the desired contradiction, and so the theorem holds.
\end{proof}









%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{A Global Linear Condition for Deadlock Freedom}
   \label{s:globCondition}
   \label{s:global.Linear}
%   \input{globLinCond}

In some cases, a simpler condition suffices to guarantee deadlock-freedom. This simpler condition is ``linear'', \ie it lacks the AND-OR alternation
aspect of $\GAO$. After execution of a reachable transition $s \la{\act} t$ of $(\B, Q_0)$, 
we consider the in-depth and out-depth of the components $\B_i \in \cmps{\act}$. There are three cases:
%
\begin{itemize}

\item \emph{Case 1}. \label{case:finite-in} $\B_i$ has finite in-depth in $\wfg{\B}{t}$: then, if $\B_i \in \SC$, it can be removed and still leave a
  supercycle $\SC'$, by Proposition~\ref{prop:supercycle:essential-subgraph-of}. Hence $\SC'$ exists in $\wfg{\B}{s}$, and so $\B_i$ is not essential
  to the creation of a supercycle.

\item \emph{Case 2}. \label{case:finite-out} $\B_i$ has finite out-depth in $\wfg{\B}{t}$: by \cor{supercycle:no-finite-outdepth}, $\B_i$ cannot be
  part of a supercycle, and so $\SC \subg \wfg{\B}{s}$.

\item \emph{Case 3}. \label{case:infinite-both} $\B_i$ has infinite in-depth and infinite out-depth in $\wfg{\B}{t}$: in this case, $\B_i$ is possibly
  an essential part of $\SC$, \ie $\SC$ was created in going from $s$ to $t$.

\end{itemize}
We thus impose a condition which guarantees that only 
Case~1 %\ref{case:finite-in}   %labels give section numbers, not list item numbers
or Case~2 %\ref{case:finite-out} 
occur. 



\begin{definition}[$\GLin(\B,Q_0,\act)$] \label{def:global:dfc}
Let $s \goesto[\act] t$ be a reachable transition of $(\B, Q_0)$.
Then, in $\wfg{\B}{t}$,  every component $\B_i$ of $\cmps{\act}$ either has finite in-depth, or has finite out-depth.
Formally, \\
\ind  $\fa \B_i \in \cmps{a}: \widepth{\B}{\B_i}{t} < \omega \lor \wodepth{\B}{\B_i}{t} < \omega.$
\end{definition}



\begin{proposition} \label{prop:indepth-finite-implies-scViol}
Assume that node $v$ of $\wfg{\B}{t}$ has a finite in-depth of $d$ in $\wfg{\B}{t}$, \ie 
$\widepth{\B}{v}{t} = d$. Then $\connViol{v}{t}$.
\end{proposition}
%
\begin{proof}
A node with finite in-depth cannot be in a wait-for-cycle, and
therefore cannot be in a strongly connected supercycle.
\end{proof}


% \prf{
% Proof is by induction on $d$.

% \noindent
% \ul{Base case, $d=1$.} Hence by Definitions~\ref{def:path} and \ref{def:depth},  
% $v$ has no incoming wait-for-edges in $\wfg{\B}{t}$. Hence by Definition~\ref{def:supercycle.violation},
% $\viol{v}{1}{t}$.

% \noindent
% \ul{Inductive step, $d > 1$.}
% By Definition~\ref{def:path} and \ref{def:depth}, all predecessors (\ie nodes $u$ such that $u \ar v \in \wfg{\B}{t}$)
% of $v$ have an in-depth of at most $d-1$. Hence, by the induction hypothesis applied to $d-1$, we obtain 
% $\viol{u}{d-1}{t}$ for all predecessors $u$ of $v$. Hence by Definition~\ref{def:supercycle.violation}, 
% Clauses~\ref{def:supercycle.violation.component.in} and \ref{def:supercycle.violation.interaction.in},
% $\viol{v}{1}{t}$.
% }




\begin{lemma} \label{lemma:glob-lin-implies-globANDOR} \label{GLinGAO}
$\fa \act \in \gamma: \GLin(\B, Q_0, \act) \imp \GAO(\B, Q_0, \act)$.  
\end{lemma}
%
\begin{proof}
Assume, for arbitrary $\act \in \gamma$, that $\GLin(\B, Q_0, \act)$ holds. That is, 
 
\ind For every reachable transition $s \goesto[\act] t$ of $(\B, Q_0)$,\\
\ind \ind  $\fas \B_i \in \cmps{\act}: \widepth{\B}{\B_i}{t} < \omega \lor \wodepth{\B}{\B_i}{t} < \omega$.

\noindent
By Propositions~\ref{prop:indepth-finite-implies-scViol} and \ref{prop:outdepth-finite-implies-scViol}, 

\ind For every reachable transition $s \goesto[\act] t$ of $(\B, Q_0)$,\\
\ind \ind  $\fas \B_i \in \cmps{\act}:  \connViol{\B_i}{t} \lor (\ex d \ge 1: \viol{\B_i}{d}{t})$.

\noindent
Hence by Definition~\ref{def:formation.violation}
and \prop{globViol-equiv-globViolDist}, 

\ind For every reachable transition $s \goesto[\act] t$ of $(\B, Q_0)$,\\
\ind \ind  $\fas \B_i \in \cmps{\act} : \formViol{\B_i}{t}$

\noindent
Hence $\GAO(\B, Q_0, \act)$ holds.
\end{proof}


\begin{theorem} \label{thm:GLin.SC-free-preserving}
$\GLin$ is supercycle-freedom preserving
\end{theorem}
%
\begin{proof}
Follows immediately from Lemma~\ref{lemma:glob-lin-implies-globANDOR} and Theorem~\ref{thm:GAO.SC-free-preserving}.
\end{proof}


%Since $\GLin(\act)$ implies $\GAO(\act)$, and $(\fa \act \in \gamma: \GAO(\act))$ implies deadlock-freedom (assuming also that initial
%state are deadlock-free), we obtain:



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Deadlock freedom using global restrictions}


\begin{corollary} [Deadlock-freedom via $\GAO, \GLin$] 
\label{theorem:global.deadlock-free}
Assume that
\bn
\item \label{theorem:global.deadlock-free.initial}
      for all $s_0 \in Q_0$, $\wfg{\B}{s_0}$ is supercycle-free, and
\item \label{theorem:global.deadlock-free.scfPres}
      for all interactions $\act$ of $\B$ (\ie $\act \in \gamma$):  $\GAO(\B, Q_0, \act) \lor \GLin(\B, Q_0, \act)$ holds.
\en
Then for every reachable state $u$ of $(\B, Q_0)$:  $\wfg{\B}{u}$ is supercycle-free, and so 
$(\B, Q_0)$ is free of local deadlock.
\end{corollary}
%
\begin{proof}
Immediate from Theorems~\ref{thm:GAO.SC-free-preserving}, \ref{thm:GLin.SC-free-preserving}, and Corollary~\ref{cor:SC-free-preserving.deadlock-free}.
\end{proof}



%% \bt[Deadlock-freedom via \GLin] 
%% \label{theorem:global:deadlock-free}
%% \label{theorem:global.linear.deadlock-free}
%% %Let $(B, Q_0)$ be a BIP system, with $B = \gamma(\B_1,\dots,\B_n)$.
%% Assume that
%% \bn
%% \item \label{theorem:global.linear.deadlock-free.initial} 
%%       for all $s_0 \in Q_0$, $\wfg{\B}{s_0}$ is supercycle-free, and
%% \item \label{theorem:global.linear.deadlock-free.globalLinear}
%%       for all interactions $\act$ of $B$ (\ie $\act \in \gamma$), $\GLin(\act)$ holds.
%% \en
%% Then for every reachable state $u$ of $(\B, Q_0)$:  $\wfg{\B}{u}$ is supercycle-free.
%% \et
%% %
%% \prf{
%% By Assumption~(\ref{theorem:global.linear.deadlock-free.globalLinear}) and 
%% Proposition~\ref{prop:glob-lin-implies-globANDOR}, we have 
%% $\fas \act \in \gamma: \GAO(\act)$. By this and
%% Assumption~(\ref{theorem:global.linear.deadlock-free.initial}), we obtain that the 
%% assumptions of Theorem~\ref{theorem:global.ANDOR.deadlock-free} hold. Hence, by applying 
%% Theorem~\ref{theorem:global.ANDOR.deadlock-free}, we obtain the conclusion

%% for every reachable state $u$ of $(\B, Q_0)$:  $\wfg{\B}{u}$ is supercycle-free.

%% and we are done.
%% }



